---
title: "ggPMX 0.8"
author: "Amine Gassem"
date: "May 3, 2018"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
library(ggPMX)
library(data.table)
knitr::opts_chunk$set(echo = TRUE)
```

## Scope 

Allo. Test.

The additional new components of the 0.8 version are : 

- Set the input parameter of the controller creator as a data.frame or data.table
- Generalize the default abbreviations tables and create set/get helper to manipulate abbreviations
- New effects global settings to rename effects variables
- Generate a single endpoint report

## Use data.frame as input 

We can now use a data.frame as an input to pmx controller. 

First we create a controller as usual and we get the input data using `get_data`
```{r}
ctr <- theophylline()
inn <- ctr %>% get_data("input")
```

We modify the input data. For example here we use pretty labels for the categorical covariates
```{r}
dat <- inn[,SEXL:=factor(SEX,levels=0:1,labels=c("Male","Female"))]

``` 

We create a new controller using the the modified data:

```{r}


  WORK_DIR <- 
    file.path( 
      system.file(package = "ggPMX"), 
      "testdata","theophylline", "Monolix")
  ctr1 <- pmx_mlx(
    config = "standing",
    directory = WORK_DIR,
    input = dat,
    dv = "Y",
    dvid = "DVID",
    cats = c("SEXL"),
    conts = c("WT0", "AGE0"),
    strats = "STUD",
    settings=pmx_settings(use.abbrev = TRUE)
  )
  
```
We can check this using : 

```{r}

ctr1 %>% pmx_plot_cats("eta_box")
```



## Use abbreviations

`ggPMX` contains a set of default abbreviations that are basically used to change the names of axes labels for one or all the plot at once.  

here an example: 
```{r}
ctr %>% pmx_plot_iwres_time(use.abbrev=TRUE)
```

To list all abbreviations we use use  `abbrev` functions. 

```{r, eval=FALSE}
abbrev()
```
But you can also the value of a single parameter : 
```{r}
abbrev("TIME")
```
We don't change those default values But we can set the value of a parameter related to a controller using :

```{r}
ctr %>% set_abbrev(TIME="NEW TIME VALUE")
```
and 
```{r}
ctr %>% get_abbrev("TIME")
```

```{r}
ctr %>% pmx_plot_iwres_time(use.abbrev=TRUE)
```

## new global settings : effects 

In order to rename effects varaibles we intrduce a new settings parameters `effects`
```{r}
ctr <- theophylline(
  settings = pmx_settings(
    effects=list(
      levels=c("ka", "V", "Cl"),
      labels=c("Concentration","Volume","Clearance")
    )
  )
)

ctr %>% pmx_plot_eta_box
ctr %>% pmx_plot_eta_hist
ctr %>% pmx_plot_eta_matrix
```

## pmx_report

`pmx_report` engine is the main enhancements of the 0.8 version. The function is well documented and contains many examples. 


1. Create a standard report using a predefined template: `standing`. This is a built-in template defined by ggPMX team. 
2. Standlone plots :  Create a standalone directory of satdnalone plots 
3. footnote feature : Plots may contain a footnote to define locations
4. Individual report; report that contain all or some of the individual plots. Buil-in template `individual`
5. Templates can be modified on the fly using the `edit` feature
6. user can define his own template as a markdwon and pass it to the `pms_report` engine


Here the simplest way to create a report using default settings: 
```{r, eval=FALSE}
ctr <- theophylline()
ctr %>% pmx_report(
  name = "my_report",
  output_type="report",
  save_dir = "some_location")
```


  