```{r load_annex, echo=FALSE,warning=FALSE,message=FALSE}
knitr::opts_chunk$set(out.width = "100%", warning = FALSE, message = FALSE)
library(ggPMX)
library(ggplot2)
library(xtable)
library(knitr)
```



## Plots table

The main target of ggPMX is to create a report containing the following plots (see abbreviation list below): 

```{r plots_list,echo=FALSE,results='asis'}

out <- rbind(
  c("Scatter plot of NPDE vs population predictions", "SCATTER", "npde_pred"),
  c("Scatter plot of NPDE vs time", "SCATTER", "npde_time"),
  c("Scatter plot of IWRES vs time", "SCATTER", "iwres_time"),
  c("Scatter plot of observations vs population predictions", "SCATTER", "dv_pred"),
  c("Scatter plot of observations vs individual predictions", "SCATTER", "dv_ipred"),
  c("Scatter plot of absolute value of IWRES vs individual predictions", "SCATTER", "abs_iwres_ipred"),
  c("Scatter plot of IWRES vs individual predictions", "SCATTER", "iwres_ipred"),
  c("Plots of observations and model predictions per individual", "IND", "individual"),
  c("Histogram of EBE", "DIS", "ebe_hist"),
  c("Boxplot of EBE", "DIS", "ebe_box"),
  c("Distribution and quantile-quantile plot of IWRES", "QQ", "qq_iwres"),
  c("Distribution and correlation structure of RE (`ETA`)", "ETA_PAIRS", "eta_matrix"),
  c("Relationships between RE and categorical covariates", "ETA_COV", "eta_cats"),
  c("Relationships between RE and continuous covariates", "ETA_COV", "eta_conts"),
  c("Visual predictive check (VPC)", "VPC", "vpc")
)

colnames(out) <- c("Plot Name", "ggPMX type", "ggPMX name")
xt <- xtable(out, label = "tab:plots_list", caption = "List of all diagnostic plots")
print(xt, comment = F)
```
Abbreviations:

- NPDE: normalized prediction distribution errors
- IWRES: individual weighted residuals
- EBE: empirical Bayes estimates
- RE: random effects
- VPC: visual predivtive check

## ggPMX main functions

`ggPMX` implements few functions to generate and manipulate diagnostic plots. 
(Should we list pmx and pmx_mlx separately and say the differences? Or it's maybe clear from the previous sections.)

```{r functions_list,echo=FALSE,results='asis'}

out <- rbind(
  c("1", "pmx, or pmx_mlx", "Creates a controller"),
  c("2", "plot_names or plots", "Lists controller plots"),
  c("3", "get_data", "Lists controller data"),
  c("4", "get_plot", "Prints a plot"),
  c("5", "set_plot", "Creates a new plot"),
  c("6", "pmx_update", "Updates an existing plot"),
  c("7", "pmx_filter", "Filters globally the data of the current session"),
  c("8", "pmx_copy", "Returns a deep copy of the controller")
)

colnames(out) <- c(" ", "Function name", "Description")

xt <- xtable(out, label = "tab:func_list", caption = "List of all `ggPMX` functions")
print(xt, comment = F)
```

(Apparently, it's not the full list. Add all functions.)
The design of the package is around the central object: the controller. It can introspected or piped using the `%>%` operand. 

**Note that**:

The controller is an `R6` object, it behaves like a reference object.  Some functions (methods) can have a side effect on the controller and modify it internally. Technically speaking we talk about chaining not piping here. However, using `pmx_copy` user can work on a copy of the controller.


## ggPMX graphical parameters 

Graphical parameters in `ggPMX` are set internally using the `pmx_gpar` function. A number of graphical parameters can be set for the different plot types. 

```{r pmx_gpar_args}
args(pmx_gpar)
```
More information can be found in the help document `?pmx_gpar` and in the examples that follow.

## Pre-defined configurations

For the moment we are mainly using standing configuration. In the next release user can specfiy configuration either by cretaing a custom yaml file or an R configuration object. Also ggPMX will create many helper functions to manipulate the configuration objects.


