---
title: "BLOQ"
author: "Amine Gassem"
date: "July 2, 2018"
output:
  pdf_document:
    toc: yes
    number_sections: yes
  html_document:
    df_print: paged
    number_sections: yes
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggPMX)
```



# Introduction

Censoring occurs when the value of a measurement or observation is only partially known. 

- The lower limit of detection **LLOD** is the lowest quantity of a substance that can be distinguished from its absence. 

- A measuring device can also have an upper limit of quantification **ULOQ** such that any value above this limit cannot be measured and reported.


We usually distinguish between three types of censoring: 

- left
- right 
- interval. 


## MONOLIX Censored observations

To define that a measurement is censored, the observation data set should include 
a CENSORING column ( default to `CENS` ) and put :

- **1** for lower limit.
- **-1** for upper limit 

Optionally, data set can contain have a limit column ( default to `LIMIT`) column to set the other limit.

## `pmx_bloq`

Whithin ggPMX, to define a bloq we use `pmx_bloq` function. 

```{r}
str(pmx_bloq(),1)
```

This will create a `pmxBLOQ` object that defines the censoring and column definition and some graphical parameters. In the following ,we will show the use of it through some examples.

# PK data below a lower limit of quantification

## Left censored data

```{r}
mlx_file <- "/home/agstudy/Downloads/2.2.censored_data/censoring1log_project.mlxtran"
ctr <- pmx_mlxtran(file_name=mlx_file)
```

```{r}
ctr %>% pmx_plot_individual(
  is.draft = FALSE,
  facets=list(ncol=3,nrow=2),
  bloq=pmx_bloq()
)
ctr %>% pmx_plot_iwres_time(bloq=pmx_bloq(color="blue"))
ctr %>% pmx_plot_dv_ipred(bloq=pmx_bloq(color="white"))

```

## Interval censored data

```{r}
mlx_file <- "/home/agstudy/Downloads/2.2.censored_data/censoring1_project.mlxtran"
ctr <- pmx_mlxtran(file_name=mlx_file)
```


```{r, echo =FALSE}
ctr %>% pmx_plot_individual(
  is.draft = FALSE,
  facets=list(ncol=3,nrow=2),
  bloq=pmx_bloq()
)

```




# PK data below a lower limit of quantification or below a limit of detection



```{r}
mlx_file <- "/home/agstudy/Downloads/2.2.censored_data/censoring2_project.mlxtran"
ctr <- pmx_mlxtran(file_name=mlx_file)

```


```{r}
ctr %>% pmx_plot_individual(
  is.draft = FALSE,
  facets=list(ncol=3,nrow=2),
  bloq=pmx_bloq()
)
```


# PK data below a lower limit of quantification and PD data above an upper limit of quantification

```{r,message=FALSE}
mlx_file <- "/home/agstudy/Downloads/2.2.censored_data/censoring3_project.mlxtran"
ctr <- pmx_mlxtran(file_name=mlx_file,endpoint="2")

```


```{r}
ctr %>% pmx_plot_individual(
  is.draft = FALSE,
  facets=list(ncol=3,nrow=2),
  bloq=pmx_bloq()
)
```


#Combination of interval censored PK and PD data
## Pk data
```{r,message=FALSE}
mlx_file <- "/home/agstudy/Downloads/2.2.censored_data/censoring4_project.mlxtran"
ctr <- pmx_mlxtran(file_name=mlx_file,endpoint="1")

```


```{r}
ctr %>% pmx_plot_individual(
  is.draft = FALSE,
  facets=list(ncol=3,nrow=2),
  bloq=pmx_bloq()
)
```

## PD data
```{r,message=FALSE}
mlx_file <- "/home/agstudy/Downloads/2.2.censored_data/censoring4_project.mlxtran"
ctr <- pmx_mlxtran(file_name=mlx_file,endpoint="2")

```


```{r}
ctr %>% pmx_plot_individual(
  is.draft = FALSE,
  facets=list(ncol=3,nrow=2),
  bloq=pmx_bloq()
)
```