---
title: "ggPMX Multiple endpoints:How it works"
author: "Amine Gassem"
date: "June 26, 2018"
output:
  html_document:
    df_print: paged
    toc: true
    theme: united
    number_sections: true
---

```{r setup, include=FALSE}
library(ggPMX)
knitr::opts_chunk$set(echo = TRUE)
```

# Context

- A model aims at predicting one (e.g., PK), two (e.g., PK and PD) endpoints. 

- The measured value of the endpoint is denoted by *DV*. *IPRED* and *PRED* are predictions for the endpoint.

- Each report is about one single endpoint. By consequence, the endpoint (if needed) is setted at the controller call `pmx` to filter the observations data set and to keep only the right values.

# ggPMX endpoint

Endpoint is defined by : 

- The name of the DVID column : `DVID` charcater value to be passed to `pmx`. can be omited if not present in the observation data.

- The endpoint value : `endpoint` can be an integer or any character value to be passed to `pmx`. `NULL` by default.


Internally , `ggPMX` will filter the obserations data set to keep only rows satisfying `DVID==endpoint`. 


# Use cases

## Case of a single endpoint (PK model)

In this case , the observation data set do not contain a `DVID` column. 
The controller will be created without passing the dvid parameter. `ggPMX` will use all the observations data set.

## Multiple endpoint (PKPD)

### Output files without prefix (first PK example)

The observation data set contains a DVID parameter with different values. 
Output files do not have a prefix ( similiar to the PK model). `ggPMX` will check the DVID values and filter data set using the given endpoint. If no endpoint is given an error will be raised to ask user to set one endpoint. 

### Output files with a prefix (other PKPD examples)

- Similiar to the previous case but `ggPMX` will use the endpoint code to  guess the output files. Basically `predictions{code}.txt` and `finegrid{code}.txt` 

The call to the `pmx` controller will be something like : 


```{r , eval = FALSE}

library(ggPMX)
pmx_mlx(
  dvid = "YTYPE", ## use this column as obseration id 
  endpoint = 1,   ## select the first endpoint 
  ...)            ## other pmx parameters , config, input,etc..
```

- ggPMX will filter observation data using YTYPE==1
- read predictions**1**.txt
- read finegrid**1**.txt

# General case:  `pmx_endpoint`



Endpoints attributes are:

- **code** (charcater :how the endpoint is coded)
- **label** (can be used in some plots title or the report file name)
- **unit**  (used in the axis label for some plots)
- **files** list of files :`list(predictions="predictions1.txt",finegrid ="finegrid1.txt")` 
- **trans**: Whether the value must be transformed before being displayed and which transformation to use.

To manage the general case , we should use the helper function `pmx_endpoint` defines as :
```{r}
ep <- pmx_endpoint(
  code = "zz",
  label = "some_label",
  unit = "some_unit",
  files = list(predictions="predictions1.txt",finegrid ="finegrid1.txt"),
  trans="log10"
)

```

Then we call the controller as :  

```{r , eval = FALSE}

library(ggPMX)
pmx_mlx(
  dvid = "YTYPE",  
  endpoint = ep,    
  ...)            
```



# Example 


```{r}

pk_pd_path <- file.path(
  system.file(package = "ggPMX"), "testdata",
  "pk_pd"
)
WORK_DIR <- file.path(pk_pd_path, "RESULTS")
ep <- pmx_endpoint(
  "4",
  files = list(
    predictions="predictions2",
    finegrid="finegrid2")
)


input_file <- file.path(pk_pd_path, "pk_pd.csv")

ctr <- pmx_mlx(
  config = "standing",
  directory = WORK_DIR,
  input = input_file,
  dv = "dv",
  dvid = "dvid",
  cats = "sex",
  conts = "wt",
  endpoint = ep
)


```

