---
title: "ggPMX - User guide"
author: "Amine Gassem"
date: "`r Sys.Date()`"
output: rmarkdown::pdf_document
vignette: >
  %\VignetteIndexEntry{ggPMX: User guide}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
\clearpage  
```{r load_package, echo=FALSE,warning=FALSE,message=FALSE}
knitr::opts_chunk$set(out.width='100%',warning=FALSE,message=FALSE)
library(ggPMX)
```



# Introduction

The `ggPMX` package generates standard model diagnostic plots and tables used in PMX activities.

The package aims to automate plot and report generation. 

This vignette introduces the ggPMX syntax, how to initialize a controller using package data , display plots, update existing plots , filter data , stratification and many others details. 

# Plots table

The main target of ggPMX is to create a report containing the following plots: 

```{r plots_list,echo=FALSE}

out <- rbind(
c("Plot of weighted residuals versus population predictions","RES","npde_pred"),
c("Plot of weighted residuals versus time","RES","npde_time"),
c("Plot of IWRES versus time","RES","iwres_time"),
c("Plot of observations versus population predictions", "RES","dv_pred"),
c("Plot of observations versus individual predictions", "RES","dv_ipred"),
c("Plot of absolute individual weighted residuals versus individual predictions","RES","abs_iwres_ipred"),
c("Plot of individual weighted residuals (IWRES) versus individual predictions","RES","iwres_ipred"),
c("Plots of observations and model predictions per individual","IND","indiv"),
c("Distribution of EBE histogram","DIS","ebe_hist"),
c("Distribution of EBE boxplot","DIS","ebe_box"),
c("Distribution and quantile-quantile plot of IWRES","QQ","qq_iwres"),
c("Distribution and correlation structure (ETA)","ECORREL","eta_matrix"),
c("Relationships between (ETA) and categorical covariates","ECORREL","eta_cats"),
c("Relationships between (ETA) and continuous covariates","ECORREL","eta_conts"),
c("Visual predictive check (VPC)","VPC","vpc"))

colnames(out) <- c("Plot Name","ggPMX type","ggPMX name")
knitr::kable(out)
```




# Package data 

In prder to esent the package features , the package comes with built in data. 
This vignette, we will use theophylline data. 
*PS: (TODO add some description of the data here, PK case,...)*


# ggpMX: basics

## theophylline controller
Within `ggPMX`, the first thing to do is to create a controller. 
To create a controller using `theophylline` data we call the helper function:

```{r theophylline_ctr}
ctr <- theophylline()
```
Let's discover the controller content by printing to the console:
```{r display ctr}
ctr
```
We get mainly 3 tables:

- Table of the controller configuration 
- Table of data sets
- Table of plots. 

**Note that**:

* We will go in depth through the controller construction details when we create a controller using custom data.
* The ctr is an object of class `r class(ctr)[1]`. It behaves like a container that stores: 
    + data configuration, 
    + plots configuration
    + data sets
    + plots
    + templates path.
    + report

## Get Plots

The controller is a container that stores all plots. To get the list of plots:

```{r plot_lists}
ctr %>% plot_names
```
Note here the use of piping 

To view a particular plot we call `get_plot` with the name of the plot:
```{r get_ipred_iwres}
ctr %>% get_plot("iwres_ipred")
```


In case of individual plots, We can pass in addition the number of page to plot to `get_plot`:

```{r ind_plots}
ctr %>% get_plot("indiv",npage = 2)
```

`get_plot` returns a standard `ggplot2` object that you can customize using any valid `ggplot2` operation. Here for example we are combining 2 versions of the distribution plot:

```{r ebe_plots, out.width='.49\\linewidth', fig.width=3, fig.height=3,fig.show='hold',fig.align='center'}
ctr %>% get_plot("ebe_hist")
ctr %>% get_plot("ebe_box")
```

## Get data

In the previous example, `ggPMX` controller reads modelling input and output data files. We can access any data set using the `get_data` function. Basically we have 2 types of data:

- **input**: Modelling input data 
- **others**: Model output data. Her we distinguish 3 or 4 files: estimates, predictions, eta and finegrid(optional)

### Modelling input data
```{r get_input_data}
ctr %>% get_data("input")
```

### Model output data
```{r get_outputs_data}
ctr %>% get_data("estimates")
ctr %>% get_data("predictions")
ctr %>% get_data("eta")
ctr %>% get_data("finegrid")
```

**Note that**

`ggPMX` uses `data.table` package to load data. Manipulating and reshaping data is very efficient and  very intuitive.  Here an example of filtering input data in a sugar syntax manner. 

```{r dx_sugar_syntax}

input <- ctr %>% get_data("input")
input[DV>350 & TIME>0  & ID==1 & SEX==1]

```



# ggPMX main functions

`ggPMX` implements few functions to generate and manipulate diagnostic plots. 

1. `pmx` , `pmx_mlx` to create a controller.
2. `plot_names`,`plots` to list controller plots.
3. `get_data` to list controller data.
3. `get_plot` to print a plot.
4. `set_plot` to create a new plot.
5. `pmx_update` to update an existing plot.
6. `pmx_filter` To filter globally session data.
7. `pmx_copy` to get a deep copy of the controller

The design of the package is around the central object the controller. that we can introspect or pipe it using the `%>%`. 

**Note that**:

The controller is an `R6` object, it behaves like a reference object.  Some functions ( methods) can have a side effect on the controller and modify it internally. Technically speaking we talk about chaining not piping here. However , using `pmx_copy` user can work on a copy of the controller.



## Appendix: All Theophylline plots

Here the list of all plots based on theophylline data set. 

### Residual plots
```{r th_plots_res, out.width='.49\\linewidth', fig.width=4, fig.height=3,fig.show='hold',fig.align='left',message=FALSE,echo=FALSE}
yy <- ctr %>% plots
yy <- as.character(unlist(yy[yy$plot_type=="RES","plot_name"]))
for(x  in yy) print(ctr %>% get_plot(x))
```
### Individual plots
```{r th_plots_ind, out.width='.49\\linewidth', fig.width=4, fig.height=3,fig.show='hold',fig.align='left',message=FALSE,echo=FALSE}
yy <- ctr %>% plots
yy <- as.character(unlist(yy[yy$plot_type=="IND","plot_name"]))
for(x  in yy) print(ctr %>% get_plot(x))
```
### Eta plots
```{r th_plots_others, out.width='.49\\linewidth', fig.width=4, fig.height=3,fig.show='hold',fig.align='left',message=FALSE,echo=FALSE}
yy <- ctr %>% plots
yy <- as.character(unlist(yy[!yy$plot_type %in% c("IND","RES"),"plot_name"]))
for(x  in yy) print(ctr %>% get_plot(x))
```
