++---
title: "ggPMX with nlmixr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ggPMX with nlmixr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
original_crayon_enabled <- getOption("crayon.enabled")
options(crayon.enabled = FALSE)
```

```{r setup}
library(ggPMX)
has_nlmixr <- FALSE
if (R.version$major >= 4) {
  library(nlmixr)
  has_nlmixr <- TRUE
}
``

It is simple to create a ggPMX controller for a nlmixr object.

Using the theophylline example with a nlmixr model we have:

```{r}
one.compartment <- function() {
    ini({
        tka <- 0.45 # Log Ka
        tcl <- 1 # Log Cl
        tv <- 3.45    # Log V
        eta.ka ~ 0.6
        eta.cl ~ 0.3
        eta.v ~ 0.1
        add.sd <- 0.7
    })
    model({
        ka <- exp(tka + eta.ka)
        cl <- exp(tcl + eta.cl)
        v <- exp(tv + eta.v)
        d/dt(depot) = -ka * depot
        d/dt(center) = ka * depot - cl / v * center
        cp = center / v
        cp ~ add(add.sd)
    })
}
if (has_nlmixr) nlmixr(one.compartment)
```

We run this model locally by:

```{r}
if (has_nlmixr) {
  fit <- nlmixr(one.compartment, theo_sd, est="saem", control=list(print=0))
}
```


The `fit` object is a nlmixr fit; You can read it into the nlmixr
controller by:

```{r}
if (has_nlmixr) {
fit %>%
    pmx_nlmixr(vpc = FALSE) -> ## VPC is turned on by default, can turn off.
    ctr ## Assigned to controller
}
```

Once the controller is created, you can of course have the same types
of diagnostic plots as the standard ggPMX package; Using the same
examples as the user manual:

```{r}
if (has_nlmixr) {
  ctr %>% pmx_plot_dv_pred
  ctr %>% pmx_plot_npde_time
  ctr %>% pmx_plot_vpc
  ctr %>% pmx_plot_eta_box
  ctr %>% pmx_plot_eta_matrix(
    shrink=pmx_shrink(size=3,hjust=1.5))
}
```

This shows the standard plots for theophylline

```{r, include = FALSE}
options(crayon.enabled = original_crayon_enabled)
```
