---
title: "How to use ggplot2 syntax with ggPMX"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{How to use ggplot syntax with ggPMX}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "100%",
  warning = FALSE,
  message = FALSE,
  fig.height=5,
  fig.width=8
)

```

We will create a simple ggPMX controller using the theophylline example:

```{r}

library(ggPMX)
library(ggplot2)

ctr <- theophylline()
```

Each ggPMX plot is an object with class ggplot:

```{r}
p <- ctr %>% pmx_plot_dv_ipred()
class(p)
```


Therefore we can use usual ggplot2 syntax to customize a ggPMX plot.

Let's look at some examples.

1. Facet

ggPMX:

```{r}
ctr %>% pmx_plot_npde_pred(strat.facet= ~SEX)
```

ggplot:

```{r}
ctr %>% pmx_plot_npde_pred() + facet_wrap(~SEX)
```

As we see in general we have the same chart, but ggPMX also automatically creates an appropriate title.
Notice, ggplot2 doesn't use the pipe %>% operator!

2. Title

ggPMX:

```{r}
ctr %>% pmx_plot_npde_pred(labels=c(title="New title!"))
```

ggplot:

```{r}
ctr %>% pmx_plot_npde_pred() + ggtitle("New title!")
```

3. X and Y axis labels

ggPMX:

```{r}
ctr %>% pmx_plot_dv_ipred(labels=c(x="X axis", y = "Y axis"))
```

ggplot:

```{r}
ctr %>% pmx_plot_dv_ipred()+xlab("X axis") + ylab("Y axis")
```