---
title: "Getting started with ggPMX"
author: "Inga Ludwig, to be added"
date: "`r Sys.Date()`"
output: html_vignette
---
<!--
%\VignetteEngine{knitr::rmarkdown}
%\VignetteIndexEntry{Getting started with ggPMX}
-->

```{r, include=FALSE}
library(ggPMX)
library(knitr)
knitr::opts_chunk$set(
    fig.width = 7,
    fig.height = 4
)

```


The `ggPMX` package has been created to facilitate easy generation of diagnostic 
plots from non-linear mixed effect model outputs.  

TODO: Add some more description here.  

Outlined below are the main steps of analyses of NLME models that are supported 
by ggPMX:

1. Reading of standard outputs from the main softwares used for NLME modelling
in Pharmatometrics (NONMEM/Monolix):

```{r, results="asis", echo=TRUE}
library("ggPMX")

# TODO: Replace the following code by some working example of how to read 
#       NONMEM/Monolix outputs

# # Calculate the sum of 1 and 2
# sum12 <- add(1, 2)
# sum12
# # Do also some subtraction
# subtract(2, 1)
```

2. Generating a full report with some standard diagnostic plots:

```{r, results="asis", echo=TRUE}
# # Retrieve the dataset 1 data item at position sum12
# retrieve_data(sum12)
# # Retrieve the dataset 2 data item at position sum12
# retrieve_data(sum12, dataset=2)
```

3. TODO: Step by step populate further sections


### References

TODO: Update the following references:  

[1] Weber S, _SciSW-R (Scientific Software in R)_. 2016; Doc.No: SciSW-R_WK_DOC001,
DiARy location: *Cabinets/Dit Asset Repository Systems/S/SciSW-R (Scientific Software in R)/SystemDevelopment & Specifications/Software Development/SciSW-R_WK_DOC001 Support Document* (also included in this package, see `inst/doc/`)  
[2] Wickham H, _R packages_. 2015; http://r-pkgs.had.co.nz/  
[3] Wickham H, _Advanced R_. 2015; http://adv-r.had.co.nz/  
[4] R Development Core Team, _Writing R Extensions_; https://cran.r-project.org/doc/manuals/r-release/R-exts.html  
[5] Collins-Sussman B et. al., _Version Control with Subversion_. 2011; http://svnbook.red-bean.com/en/1.6/index.html  

### R Session Info

```{r}
# Finally, it is good practice to include session info
# information in the vignette:
sessionInfo()
date()
```

