---
output:
  pdf_document: default
  html_document: default
---
```{r load_intro, echo=FALSE,warning=FALSE,message=FALSE}
knitr::opts_chunk$set(out.width = "100%", warning = FALSE, message = FALSE)
library(ggPMX)
library(ggplot2)
library(xtable)
library(knitr)

theophylline <- file.path(system.file(package = "ggPMX"), "testdata", "theophylline")
work_dir <- file.path(theophylline, "Monolix")
input_data <- file.path(theophylline, "data_pk.csv")

ctr <- theophylline()
```


The `ggPMX` package generates standard diagnostic plots and tables for mixed effect models used in Pharmacometric (PMX) activities. The tool is built upon the ggplot2 package and is planned to support models developped either with Monolix or NONMEM software.
(The current release (0.6.1) only supports Monolix models.)

The package aims to provide a workflow that is consistent, efficient and which results in high quality graphics ready to use in official documents and reports. The package allows a high degree of flexibility and customization, yet providing an acceptable default setting. The package will also allow to fully automate plots and report generation (in a next release). 

The general context is the analysis of mixed effect models fitted to data. ggPMX was developed in the framework of Pharmacometric activities, in which case the model is a population pharmacokinetic (PK) and/or pharmacodynamic (PD) model and the data is clinical or pre-clinical PK and/or PD data.

In the context of model building, evaluation and qualification, it is good practice to assess the goodness-of-fit of models by inspecting (qualitatively and quantitatively) a set of graphs that indicate how well the model describes the data. Several types of diagnostic plots allow to evaluate a mixed effects model fit, the most common being:

a. residual-based plots
b. prediction-based plots
c. empirical Bayes estimates (EBE)-based plots
d. simulation-based plots.

The following figures are examples of diagnotic plots. 
```{r illustrate_diagnostic, out.width='.25\\linewidth', fig.width=4, fig.height=4, fig.show='hold', fig.align='center', echo=FALSE}
ctr <- theophylline()
ctr %>% pmx_plot_dv_pred
ctr %>% pmx_plot_eta_box
ctr %>% pmx_plot_eta_hist(facets=list(scales="free_y"))
ctr %>% pmx_plot_individual(npage = 1,
                            is.legend=FALSE,
                            facets=list(ncol=2,nrow=3))
ctr %>% pmx_plot_eta_matrix(
  shrink=list(size=3,hjust=1.5))
```


This vignette introduces the ggPMX syntax. Section X describes the general architecture and defines key design elements (the controller and several data frames). Section X+1 describes how to initialize a controller using a built-in model. This section allows to have a quick grasp of the package capabilities by showing how to display plots, update existing ones, how to filter and stratify data. Section X+2 explains how to use the package with custom models. 

