---
output:
  pdf_document: default
  html_document: default
---
```{r load_settings, echo=FALSE,warning=FALSE,message=FALSE}
knitr::opts_chunk$set(out.width = "100%", warning = FALSE, message = FALSE)
library(ggPMX)
library(ggplot2)
library(xtable)
library(knitr)

theophylline <- file.path(system.file(package = "ggPMX"), "testdata", "theophylline")
work_dir <- file.path(theophylline, "Monolix")
input_data <- file.path(theophylline, "data_pk.csv")

ctr <- theophylline()
```

User can define controller with global settings that we applied to all plots. For example remove draft annoataion , use abbreviation defintions to define axis labels , etc,. 

## pmx_settings

A settings object is defined by `pmx_settings`. The created object is passed as a parameter to `pmx` to initialize it globally.

see `?pmx_settings`  to get the list of global settings with their corresponding default values. Here the general syntax:

```{r settings_example, fig.width=5, fig.height=4,eval=FALSE}

## set one or more settings
my_settings <- pmx_settings(
  is.draft = FALSE,
  use_abbrev=TRUE,
  ...) ### set other settings parameters here 
ctr <-
  pmx_mlx(
    ..., ## put here other pmx parametes 
    settings = my_settings
  ) 
```


## remove draft label globally 

The standing configuration defines all the plot as a draft version. You can create an official version of all plots using `is.draft=FALSE`:

```{r settings_is_draft,fig.height=3, fig.width=3, fig.show='hold', fig.align='center'}

ctr <- theophylline(settings=pmx_settings(is.draft = FALSE))

ctr %>% pmx_plot_dv_pred
ctr %>% pmx_plot_ebe_box


```


## Use abbreviation definitions 

The standing configuration initialize all the plot using abbreviations for axis labels. Each abbreviation has its corresponding definition. To get the list of abbreviation : 

```{r settings_get_abbrev}
ctr %>% get_abbrev
```
You can update one abbreviation to set a custom label 

```{r settings_set_abbrev}
ctr %>% set_abbrev(TIME="TIME after the first dose")
```

Using `use_abbrev` flag you can use abbreviation definition to set axis labels:

```{r settings_use_abbrev}
ctr <- theophylline(settings=pmx_settings(use_abbrev = TRUE))
ctr %>% set_abbrev(TIME="Custom TIME axis")
ctr %>% pmx_plot_npde_time


```

## Use finegrid file 

within Monolix, user can choose to not use finegrid file even if it is present. 

```{r settings_use_finegrid}
ctr <- theophylline()
ctr <- pmx_plot_individual(use.finegrid =FALSE)


```

## Set strat color legend

In case of color startfication user can customize the legend. For example here using the `ggplot2::scale_color_manual`:

```{r settings_color_scales_local}
ctr <- theophylline()
ctr %>% pmx_plot_npde_time(strat.color="STUD")+ 
      ggplot2::scale_color_manual(
        "Study",
        labels=c("Study 1","Study 2"),
        values=c("1"="lightyellow","2"="lightblue"))
    

```

Another way to do it is to define a global `scales.color` parameter that will applied to all plots with strat.color :
```{r serrings_solor_scales}
ctr <- theophylline(
  settings=
    pmx_settings(
      color.scales=list(
        "Study",
        labels=c("Study 1","Study 2"),
        values=c("1"="lightyellow","2"="lightblue"))
    )
)

ctr %>% pmx_plot_npde_time(strat.color="STUD")

ctr  %>%  pmx_plot_ebe_box(strat.color="STUD")
```


## Define categorical labels :

In case of faceting by stratification user can redfine categorical labels to have more human readables strips. Lables are defined within `cats.labels` argument and user can use them by setting `use_lables` to TRUE.

```{r settings_cat_labels}


ctr <- theophylline(
  settings=
    pmx_settings(
      cats.labels=list(
        SEX=c("0"="M","1"="F"),
        STUD=c("1"="Study 1","2"="Study 2")
      ),
      use_labels = TRUE
    )
)


ctr %>%   pmx_plot_npde_time(strat.facet=~SEX)
ctr  %>%  pmx_plot_ebe_box(strat.facet =~SEX)

```


