```{r load_shrink, echo=FALSE,warning=FALSE,message=FALSE}
knitr::opts_chunk$set(out.width = "100%", warning = FALSE, message = FALSE)
library(ggPMX)
library(ggplot2)
library(xtable)
library(knitr)

theophylline <- file.path(system.file(package = "ggPMX"), "testdata", "theophylline")
work_dir <- file.path(theophylline, "Monolix")
input_data <- file.path(theophylline, "data_pk.csv")

ctr <- theophylline()
```

## Default call

The shrinkage is a computation within controller data. In general it is used to annotate the plots. Although one can get it independently from any plot using `pmx_comp_shrink`. It is part of the `pmx_compt_xx` layer( In the future we will add , `pmx_comp_cor` , `pmx_comp_summary`,..)

Here the basic call:

```{r shrink_comp}
ctr %>% pmx_comp_shrink
```


We get the shrinkage for each effect (**SHRINK** column).

The same values can be shown on distribution plot , for example  : 

```{r shrink_plot_box}
ctr %>% pmx_plot_ebe_box

```
or : 
```{r shrink_plot_hist}
ctr %>% pmx_plot_ebe_hist

```

You can add or remove shrinkage annotation using `has.shrink` argument ( TRUE by default) : 

```{r shrink_plot_no}
ctr %>%   pmx_plot_ebe_box( has.shrink = FALSE) 

```



## Var function

You can compute shrinkage by applying either standard deviation `sd` or variance `var` : 
```{r, compute_var }
ctr %>% pmx_comp_shrink(  fun = "var")
```

```{r, shrink_plot_var}
ctr %>% pmx_plot_ebe_box( shrink=list(fun = "var"))
```
 
 
## Shrinkage and stratification

Shrinkage can be applied after stratification :

```{r shrink_comp_strat}
ctr %>% pmx_comp_shrink(strat.facet = ~SEX)
```
or by grouping like : 

```{r shrink_comp_strat_color}
ctr %>% pmx_comp_shrink(strat.color = "SEX")
```

We can 
```{r shrink_plot_strat, fig.width=9, fig.height=8}
ctr %>% pmx_plot_ebe_hist(has.shrink = TRUE, strat.facet = ~SEX,
                          facets=list(scales="free_y"))
```
or 
 
```{r fig.width=9, fig.height=8}
ctr %>% pmx_plot_ebe_box(has.shrink = TRUE, strat.facet = "SEX",
                          facets=list(scales="free_y",ncol=2))
```

