---
title: "PMX configuration objects"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

In this article we provide a tutorial on the structure of PMX configuration objects. It is intended primarily for developers interested in understanding the internal workings of the package.

```{r setup}
library(ggPMX)
```

First, let's construct a PMX controller object manually, using the Monolix test data files supplied with the package. 

```{r}
# paths
ggpmx_dir        <- system.file(package = "ggPMX")
theophylline_dir <- file.path(ggpmx_dir, "testdata", "theophylline")
work_dir         <- file.path(theophylline_dir, "Monolix")
input_file       <- file.path(theophylline_dir, "data_pk.csv")

# identical to the example in the pmx() docs
ctr <- pmx(
  sys = "mlx",
  config = "standing",
  directory = work_dir,
  input = input_file,
  dv = "Y",
  dvid ="DVID"
)
ctr
```

This article unpacks the `config` argument, showing the key data structures involved, and some notes that may assist when writing pull requests.

## PMX configuration object

The key role played by the `config` argument to `pmx()` is to construct the data with the (public) `$config` field of the controller object. As you might expect from the name, this is a data structure that stores configuration information about the plots. When printed, it produces this:

```{r}
ctr_config <- ctr$config
ctr_config
```

The object itself is has S3 class `pmxConfig`: 

```{r}
class(ctr_config)
```

This object stores several kinds of configuration:

```{r}
names(ctr_config)
```

## Constructing a config object

When calling `pmx()` to construct a controller, you *could* choose to pass a `pmxConfig` object directly, but it's not required. If you look at the source code for the `pmx()` function, you can see that if the user doesn't specify a `config` argument, ggPMX will apply the default "standing" configuration, and then loads it using the `load_config()` function before passing the configuration object to the initialiser function `pmxClass$new()`. 

Here's what happens when we call `load_config()` directly:

```{r}
standing_config <- load_config("standing", sys = "mlx")
standing_config
```

This is the configuration object that gets stored within the PMX controller.

## YAML configuration files

The next "layer" down in unpacking the configuration object is understanding where the configuration information is stored. If we look at the source code for `load_config`, we can see that it expects to find two configuration files: there is a "plot PMX" file that has a `.ppmx` file extension, and an "input PMX" file with a `.ipmx` extension:

```{r}
load_config
```

These are both YAML files. The paths to the files for the default "standing" configuration are as follows: 

```{r}
plots_template  <- file.path(ggpmx_dir, "init", "standing.ppmx")
inputs_template <- file.path(ggpmx_dir, "init", "templates", "mlx", "standing.ipmx")
```

Internally the `load_config()` function passes the work to `ggPMX:::load_config_files()`, which in turn uses `yaml::yaml.load_file()` to parse the YAML file and return a list. Just to illustrate what it looks like, we'll read the raw plots file without any parsing and look at what it contains. The file specifies a list of plot names, and default argument values that should be applied when that plot is built (unless of course the user overrides them). 

Here's what it looks like:

```{r}
plots_config_raw <- brio::read_lines(plots_template)
cat(plots_config_raw[1:32], sep = "\n") # show first 32 lines only
```

Here's what that same config looks like after YAML parsing:

```{r}
plots_config <- yaml::yaml.load_file(plots_template)
str(plots_config[1]) # show ABS_IWRES_IPRED only
```

From the developer perspective, this "standing" configuration represents a default that ggPMX will apply to most plots. If you introduce a new plot (or modify an existing one), you may need to modify the `standing.ppmx` file so that those changes are applied by default. 

## Custom configuration

It's not as relevant from the developer perspective, but worth noting that ggPMX also provides user facing functions to define their own configuration using `pmx_config()`. The package supplies an example of this:

```{r}
custom_config <- pmx_config(
  plots = file.path(ggpmx_dir, "examples", "plots.yaml"),
  inputs = file.path(ggpmx_dir, "examples", "custom_inputs.yaml")
)
custom_config
```

Having done this, you could pass the configuration object to `pmx()`, like this:

```{r}
ctr2 <- pmx(
  sys = "mlx",
  config = custom_config,
  directory = work_dir,
  input = input_file,
  dv = "Y",
  dvid ="DVID"
)
```

As above. When changing a plot it may be necessary to alter the "examples" configuration files.

